Region-network hierarchical sparsity priors for\\ high-dimensional inference in brain imaging

### 2016-01-27 paper/paper.tex
typo
rationale

% SPECIAL
Functional specialization has been
explored and interpreted based on many different research methods.
%
Single cell recordings and microscopic examination
revealed, for instance, the
specialization in the occipital visual cortex into V1, V2, V3, V3A/B, and V4
\cite{hubel1962receptive, zeki1978functional}.
Tissue lesion of the mid-fusiform gyrus of the visual system,
for instance,
was frequently reported to impair
recognition of others' identity from faces
\cite{iaria2008contrib}.
The whole-brain localization of
sensory, motor, and emotional functions to cortical areas
% in the living brain
was later enabled by
non-invasive brain imaging with
functional magnetic resonance imaging (fMRI) and
positron emission tomography (PET)
\cite{fristen1997imaging}.
Further,
radioactive mapping of neurotransmitter receptors
rendered accessible yet another
local characteristic of neuronal populations
\cite{zilles2009receptor}.
% allude to our many-regions layer in our methods
In the computational era,
automatic clustering methods are increasingly employed to
regionally differentiate the cerebral cortex,
which can partly be more fine-grained than
classical microscopical borders
\cite{behrens03, cbp2015review}.
Today,
high-thoughput approaches enable
ultrahigh-resolution 3D models of brain anatomy
at near-cellular scale
\cite{amunts2013bigbrain}.
%
As a crucial common point,
all these methodological approaches
yield neuroscientific findings
that are naturally interpreted according to
non-overlapping, discrete region compartments
as the basic architecture of brain organization.

@@

\section{Methods}
We make use of the hierarchical group lasso as introduced by Jennaton and 
others in \cite{jenatton2011structured} to create a set of convex penalties
with which we can probe the interplay between regional specificity and 
inter-regional functional integration during different cognitive tasks.
%
\paragraph{Rationale}
Th

@@

\paragraph{Benchmarking hierarchical tree sparsity against
common sparsity penalties.}
Hierarchical region-network priors have been systematically
evaluated against other popular choices of sparse classification algorithms
in an 18-class scenario
(Figure \ref{fig_sparsities}).
%
Logistic regression with $\ell_1/\ell_2$ block norm penalization
incorporated a hierarchy of previously known region and network neighborhoods
for a neurobiological bias of the statistical estimation.
%
Vanilla logistic regression with $\ell_1$-penalization
does not assume any previously known special structure.
This classification estimator embraces a vision of neural activity structure
that expectes a minimum of
topographically and functionally independent brain voxel to be relevant.
%
Logistic regression with sparse group sparsity
imposes a structured $\ell_1/\ell_2$ block norm with additional $\ell_1$ term
with a known atlas of region voxel groups onto the statistical estimation process.
This supervised estimator shrinks and selects the coefficients
of topographically compact voxel groups expected to be relevant together.
%
Logistic regression with trace-norm penalization
imposed low-rank structure.
This supervised classification algorithm
expected a minimum of unknown network patterns
to be relevant.
%
The stratified and shuffled training data were submitted to
to a nested cross-validation scheme
for model selection and model assessment.
In the inner CV layer, the logistic regression estimators
have been trained in a one-versus-rest design that
distinguishes each class from
the respective 17 other classes
(number of maximal iterations=$100$, tolerance=$0.001$).
In the outer CV layer, grid search
selected among candidates for the respective $\lambda$ parameter
by searching between $10^{-2}$ and $10^{1}$ in 9 steps on a logarithmic scale.
Importantly, the thus selected sparse logistic regression classifier was
evaluated on an identical test set in all settings.
%
Hierarchial tree sparsity demonstrated the best generalization
in distinguihsing unseen neural activity maps from 18 tasks
(89.67\%, mean recall XX.YY, mean precision XX.YY).
It was closely followed by logistic regression with trace-norm
regularization that is structured but not neurobiologically informed
(89.43\%, mean recall XX.YY, mean precision XX.YY).
Lasso featured an average performance comparing to the other sparse estimators
(88.60\%, mean recall XX.YY, mean precision XX.YY)
Introduced a priori knowledge of brain region compartments
by sparse group sparsity performed worst
(87.89\%, mean recall XX.YY, mean precision XX.YY).
%
In sum,
biasing sparse model selection by domain knowledge of region-network hierarchies
outperformed other types of frequently used sparse penalization techniques.

>>>>

% SPECIAL
Functional specialization has been
explored and interpreted based on many different research methods.
%
Single cell recordings and microscopic examination
revealed, for instance, the
specialization in the occipital visual cortex into V1, V2, V3, V3A/B, and V4
\cite{hubel1962receptive, zeki1978functional}.
Tissue lesion of the mid-fusiform gyrus of the visual system,
for instance,
was frequently reported to impair
recognition of others' identity from faces
\cite{iaria2008contrib}.
The whole-brain localization of
sensory, motor, and emotional functions to cortical areas
% in the living brain
was later enabled by
non-invasive brain imaging with
functional magnetic resonance imaging (fMRI) and
positron emission tomography (PET)
\cite{fristen1997imaging}.
Further,
radioactive mapping {\color{red} Michael: is this the correct term?} of neurotransmitter receptors
rendered accessible yet another
local characteristic of neuronal populations
\cite{zilles2009receptor}.
% allude to our many-regions layer in our methods
In the computational era,
automatic clustering methods are increasingly employed to
regionally differentiate the cerebral cortex,
which can partly be more fine-grained than
classical microscopical borders
\cite{behrens03, cbp2015review}.
Today,
high-throughput approaches enable
ultrahigh-resolution 3D models of brain anatomy
at near-cellular scale
\cite{amunts2013bigbrain}.
%
As a crucial common point,
all these methodological approaches
yield neuroscientific findings
that are naturally interpreted according to
non-overlapping, discrete region compartments
as the basic architecture of brain organization.

@@

\section{Methods}
Our main contribution is the domain-specific adaptation of sparse structured
penalties making it possible to jointly incorporate functional 
specialization and functional integration priors into statistical 
estimators.
We make use of the hierarchical group lasso as introduced by Jennaton and 
others in \cite{jenatton2011structured} to create a set of convex penalties
with which we can probe this interplay between regional specificity and 
inter-regional functional integration during different cognitive tasks.
%
\paragraph{Rationale}
Three-dimensional voxel brain images as obtained by various neuroimaging
techniques are very high-dimensional, but also very highly structured.
While its explicit dimensionality, the number of voxels, which varies with
the resolution of the image is generally of the order of 100000, the number
of samples available for analyses never exceeds hundreds or thousands.
This \(n << p\) situation immediately implies an underdeterminedness for any
linear model basing its decisions on dot products with the voxel values.

Luckily, there is and abundance structure to be exploited, which can be
injected into the estimation as a domain knowledge prior. Here we are 
interested in incorporating knowledge about both the well-established 
modules of functional specialization and the interactions of these modules
amongst each other across the brain.
%
Developmentally, such large-scale networks emerge during late
fetal growth (Doria et al., 2010), before cognitive capacities mature
in childhood. 
%
In adults, nodes of a same cohesive network have more
similar functional profiles than nodes from different networks
(Anderson et al., 2013).
%
On the signal level, functional modules are localized structures implicating
neighboring voxels in an image. A variety of known functional modules has
been compiled into atlases of the brain. Different types of atlases rely on
different ways of segregating modules from another. For instance, anatomical
atlases rely on the variation of tissue properties across the brain, which
can be at the cellular level or based on higher-order structure.
%
Global functional networks can be discovered by statistical analysis of
functional brain imaging data, for example the BOLD contrast in fMRI. A
well-established method of estimating these networks is independent 
component analysis \cite{Hamalainen}. It robustly identifies the default
mode, saliency and attention networks.
%
Since these network maps are for the most part disjoint spatially

@@

\paragraph{Benchmarking hierarchical tree sparsity against
common sparsity penalties.}
Hierarchical region-network priors have been systematically
evaluated against other popular choices of sparse classification algorithms
in an 18-class scenario
(Figure \ref{fig_sparsities}).
%
Logistic regression with $\ell_1/\ell_2$ block norm penalization
incorporated a hierarchy of previously known region and network neighborhoods
for a neurobiological bias of the statistical estimation.
%
Vanilla logistic regression with $\ell_1$-penalization
does not assume any previously known special structure.
This classification estimator embraces a vision of neural activity structure
that expects a minimum of
topographically and functionally independent brain voxel to be relevant.
%
Logistic regression with sparse group sparsity
imposes a structured $\ell_1/\ell_2$ block norm with additional $\ell_1$ term
with a known atlas of region voxel groups onto the statistical estimation process.
This supervised estimator shrinks and selects the coefficients
of topographically compact voxel groups expected to be relevant together.
%
Logistic regression with trace-norm penalization
imposed low-rank structure.
This supervised classification algorithm
expected a minimum of unknown network patterns
to be relevant.
%
The stratified and shuffled training data were submitted to
to a nested cross-validation scheme
for model selection and model assessment.
In the inner CV layer, the logistic regression estimators
have been trained in a one-versus-rest design that
distinguishes each class from
the respective 17 other classes
(number of maximal iterations=$100$, tolerance=$0.001$).
In the outer CV layer, grid search
selected among candidates for the respective $\lambda$ parameter
by searching between $10^{-2}$ and $10^{1}$ in 9 steps on a logarithmic scale.
Importantly, the thus selected sparse logistic regression classifier was
evaluated on an identical test set in all settings.
%
Hierarchial tree sparsity demonstrated the best generalization
in distinguishing unseen neural activity maps from 18 tasks
(89.67\%, mean recall XX.YY, mean precision XX.YY).
It was closely followed by logistic regression with trace-norm
regularization that is structured but not neurobiologically informed
(89.43\%, mean recall XX.YY, mean precision XX.YY).
Lasso featured an average performance comparing to the other sparse estimators
(88.60\%, mean recall XX.YY, mean precision XX.YY)
Introduced a priori knowledge of brain region compartments
by sparse group sparsity performed worst
(87.89\%, mean recall XX.YY, mean precision XX.YY).
%
In sum,
biasing sparse model selection by domain knowledge of region-network hierarchies
outperformed other types of frequently used sparse penalization techniques.
